Online EV Scheduling Algorithms for Adaptive Charging Networks with Global Peak Constraints

### 2019-08-27 TSC/R1/TSC.tex
MH update 2

$\vartriangleright$ We conduct a set of  simulations to evaluate the performance of our proposed algorithms. 
%For offline scheduling problem under integral charging model, the results  demonstrate that the proposed approximation algorithm is close to the optimal solution and is significantly better than the obtained theoretical bounds. 
The results of online algorithms for both integral and fractional settings are close to the optimum (within $90\%$ and $94\%$ for integral and fractional models in a representative scenario). In addition, our algorithm outperforms the existing scheduling algorithm in Caltech ACN~\cite{lee2016adaptive} by $35\%$ for integral revenue model.
%The simulation section also discusses on the positive and negative effects of slackness parameter in different scenarios.

@@

\section{Related Work}
\label{sec:rel}
\revv{There is an extensive work in the general topic of EV demand management and scheduling~\cite{mukherjee2015review} by considering different scenarios such as optimal operations with EV coordinations~\cite{mukherjee2017distributed}, EV scheduling with incorporation of renewable energy, and energy storage systems~\cite{de2017impact,shafie2018innovative}. In this section, we focus on the literature related to peak-constrained EV charging scheduling in single and multiple stations.}
\subsubsection{Peak-Constrained EV Charging Scheduling}
\label{sec:rel:peakconstrained}			
\revv{There is an extensive literature on EV charging scheduling problem focusing on single CS~\cite{Tang,Wen}, while the local and global peak constraints are omitted or only the local peak is considered.}	As we discuss in Section~\ref{sec:problem}, the global optimal solution cannot be obtained by separately solving the single station problems. Hence, those solutions cannot be directly applied to the multiple CS scenario with global peak constraints.

@@

%\cite{Shroff2014} and \cite{Stein2012} study social welfare maximization problem while the profit of both EV owners and CS is considered.  social welfare of the EV owners \cite{Stein2012}, charging cost \cite{Tang,WTang}, and total value of fully charged EVs \cite{Azar}). In addition, \cite{WTang, Tang,Shroff2014,Zhao} do not take into account the peak limitation of CS. }
		\section{System Model and Problem Formulation}
		\label{sec:model}

@@

The proposed algorithms compared to the optimal offline solution of integral model ($\textsc{iOPT}$) and $\iolp$.%%% which is the running online algorithm in the Caltech ACN \cite{lee2016adaptive}. 
%The proposed algorithms for fractional revenue model, \fcs and \focs, are compared to fractional version of OLP algorithm (referred to as \folp), which is implemented in Caltech ACN  (see~\cite{alinia2017online} for details). 
Recall that \fcs is optimal offline solution. The notable observations  are  as  follows:  (i)  the  general  trend in scenario of Fig.~\ref{fig:V-N-M} is that  by increasing  the  number  of  EVs,  total  revenue  increases.  This is because with more number of EVs, the scheduler has more freedom to choose more valuable EVs. (ii)  as  explained in Section~\ref{sec:fractional}, under fractional charging model, better results are expected due to increased scheduling flexibility in CSs. According to the simulation data that we extracted from Fig.~\ref{fig:V-N-M}, the gain obtained by \fcs and \focs in Fig. \ref{fig:V-N-M} are respectively \revv{$12\%$ and $10\%$} more than the gain of the \ics and \iocs. (iii) \revv{In Fig. \ref{fig:V-N-M2} and Fig. \ref{fig:V-N-M4}, sum of the local peaks is less than or equal to the global peak while in Fig. \ref{fig:V-N-M8} this sum is two times greater than the global peak. Consequently, the total revenue significantly improved when the number of CSs is increased from $2$ to $4$ while there is a slight improvement from $4$ to $8$ CSs as the global peak constraint prevents the algorithms from charging more EVs.}
 (iv) in integral revenue model, the proposed \iocs algorithm acts significantly better than \iolp. In particular, \iocs improves \iolp by \revv{$38\%, 36\%$, and $32\%$} for $m=2$, $m=4$ and $m=8$, respectively. \revv{In fractional revenue model, however, there is a very slight difference between \focs and \folp.}
 %We highlight that \iocs (\focs) is also better choice in terms of the algorithm complexity compared to \iolp (\folp).
(v) \ics approximates \textsc{iOpt} by \revv{$97\%$, $94\%$, and $95\%$} for $m=2$, $m=4$, and $m=8$, respectively. On average, \iocs is \revv{$90\%$} of \textsc{iOpt} and, \focs is \revv{$94\%$} of its optimal offline solution, \fcs.  
(vi) finally, the results depict that \ics achieves much better results in practice as compared to the theoretical approximation ratio that characterizes the performance in worst-case scenario.

>>>>

$\vartriangleright$ We conduct a set of  simulations to evaluate the performance of our proposed algorithms. 
%For offline scheduling problem under integral charging model, the results  demonstrate that the proposed approximation algorithm is close to the optimal solution and is significantly better than the obtained theoretical bounds. 
The results of online algorithms for both integral and fractional settings are close to the optimum (within $90\%$ and $94\%$ for integral and fractional models in a representative scenario). In addition, our algorithm outperforms the existing scheduling algorithm in Caltech ACN~\cite{lee2016adaptive} by $35\%$ for integral revenue model.

\bt{This paper represents follow-on work to our previous work~\cite{alinia2018competitive}, where we study a simplified version of the problems studied in this paper in fractional revenue model. The competitive analysis in~\cite{alinia2018competitive} provide a competitive ratio of different algorithms when all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, extend the results and propose a 2-competitive online algorithm with no assumption on the input parameters. Also, the current study investigates both fractional and integral revenue models while considering global peak constraint in the formulation and addressing multiple charging station scenario. Last, we consider more realistic trace-driven experiments in this paper and compare the results to an existing real-world scheduling algorithm.}
%The simulation section also discusses on the positive and negative effects of slackness parameter in different scenarios.

@@

\section{Related Work}
\label{sec:rel}
\revv{There is an extensive work in the general topic of EV demand management and scheduling~\cite{mukherjee2015review} by considering different scenarios such as optimal operations with EV coordination and congestion management~\cite{mukherjee2017distributed,rigas2013congestion,hu2013coordinated}, EV scheduling with incorporation of renewable energy, and energy storage systems~\cite{de2017impact,shafie2018innovative}, and pricing and bidding~\cite{huang2014distribution,hou2019bidding} In this section, we focus on the literature related to peak-constrained EV charging scheduling in single and multiple stations.}
\subsubsection{Peak-Constrained EV Charging Scheduling}
\label{sec:rel:peakconstrained}			
\revv{There is an extensive literature on EV charging scheduling problem focusing on single CS~\cite{Tang,Wen}, while the local and global peak constraints are omitted or only the local peak is considered.}	As we discuss in Section~\ref{sec:problem}, the global optimal solution cannot be obtained by separately solving the single station problems. Hence, those solutions cannot be directly applied to the multiple CS scenario with global peak constraints.

@@

%\cite{Shroff2014} and \cite{Stein2012} study social welfare maximization problem while the profit of both EV owners and CS is considered.  social welfare of the EV owners \cite{Stein2012}, charging cost \cite{Tang,WTang}, and total value of fully charged EVs \cite{Azar}). In addition, \cite{WTang, Tang,Shroff2014,Zhao} do not take into account the peak limitation of CS. }

\subsubsection{Worst-case Analysis in Similar Scheduling Problems}
\label{sec:rel:comp}
\bt{
	Similar underlying scheduling problems with slightly different settings has been studied in the literature in both offline and online settings. In offline setting, the problem is more interested in integral revenue model where the problem becomes a combinatorial optimization problem and approximation algorithms have been used to find effective solutions. 
The performance of an approximation algorithm is determined by its \emph{approximation ratio} for offline 
algorithms. On the other hand, competitive online algorithms are used in the online setting and \emph{competitive ratio} is the performance metric which compare the algorithm's result to the offline optimal solution. 
Under integral revenue model, \cite{jain2015near} proposes an offline algorithm for scheduling of batch jobs in cloud computing which is similar to EV scheduling problem. It is assumed that all jobs are available to be processed at time 0. The authors propose a $\frac{C}{C-k}.\frac{s}{s-1}$-approximation algorithm where $C$ is the cloud capacity and $s$ is a ``slackness'' parameter. 
Similarly, \cite{yao2016real} tackles the offline integral problem and proposes a convex relaxation method to find a near-optimal solution. No theoretical bound is provided for the algorithm and the performance is examined by simulation results. In this paper, we propose \ics for the EV scheduling problem with slightly different settings in the constraint sets, and tackle the problem in online setting in which the full knowledge of the online parameters is not available a priori, and provide competitive ratio under specific scenarios. 
In case of fractional revenue model (for offline case), the scheduling problem is more straightforward to tackle since the underlying problem is linear. In this paper, we propose \fcs as an optimal solution for offline fractional case.

In the online scenario, the scheduling problem is fundamentally more challenging. In our recent study \cite{alinia2018competitive}, 
we proposed two online algorithms referred to as WFAIR and WRAND for EV scheduling problem. The proposed algorithms provide a competitive ratio of $2-1/U$ where $U$ is "scarcity level". However, \cite{alinia2018competitive} the result only holds when all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, we propose \focs as a 2-competitive online algorithm with no assumption on the input parameters. Also, the current study investigates both fractional and integral revenue model while considering total peak constraint in the formulation and addressing multiple charging station scenario. 
In another work~\cite{firstfit}, two simple and natural online algorithms called \textsc{FirstFit} and \textsc{EndFit} were developed and are proved to be 2-competitive. \focs is extension of these algorithms by taking into account the maximum charging rate of the EVs and multiple charging station scenario (see ``Remarks'' at Section IV). As recent studies, \cite{zheng2016online,deng2016whether} provide online algorithms for EV scheduling problem, however, they do not take into account global peak constraint in the underlying problem.

Another direction is to tackle online scenario under integral revenue model. In this setting, the scheduling problem becomes fundamentally more challenging. We note that the integral scheduling problem is strongly NP-Hard even in the offline case \cite{de2018complexity}. Due to combined online and combinatorial challenges, there are very limited studies in this category. \cite{lucier2013efficient} provides an online algorithm without considering the maximum processing rate of the jobs where its competitive ratio can be arbitrary bad depending on ``slackness'' parameter'. Our work in Section~\ref{sec:integral} also takes into account a notion of slackness parameter in our proposed online solution, and in addition, considers \iocs for multiple station scenario and maximum charging rate.
}

\section{System Model and Problem Formulation}
		\label{sec:model}

@@

The proposed algorithms compared to the optimal offline solution of integral model ($\textsc{iOPT}$) and $\iolp$. \bt{Note that comparison with offline optimal could be considered as a baseline comparison with the category of approximate offline algorithms in integral model~\cite{yao2016real}.}

%%% which is the running online algorithm in the Caltech ACN \cite{lee2016adaptive}. 
%The proposed algorithms for fractional revenue model, \fcs and \focs, are compared to fractional version of OLP algorithm (referred to as \folp), which is implemented in Caltech ACN  (see~\cite{alinia2017online} for details). 
Recall that \fcs is optimal offline solution. The notable observations  are  as  follows:  (i)  the  general  trend in scenario of Fig.~\ref{fig:V-N-M} is that  by increasing  the  number  of  EVs,  total  revenue  increases.  This is because with more number of EVs, the scheduler has more freedom to choose more valuable EVs. (ii)  as  explained in Section~\ref{sec:fractional}, under fractional charging model, better results are expected due to increased scheduling flexibility in CSs. According to the simulation data that we extracted from Fig.~\ref{fig:V-N-M}, the gain obtained by \fcs and \focs in Fig. \ref{fig:V-N-M} are respectively \revv{$12\%$ and $10\%$} more than the gain of the \ics and \iocs. (iii) \revv{In Fig. \ref{fig:V-N-M2} and Fig. \ref{fig:V-N-M4}, sum of the local peaks is less than or equal to the global peak while in Fig. \ref{fig:V-N-M8} this sum is two times greater than the global peak. Consequently, the total revenue significantly improved when the number of CSs is increased from $2$ to $4$ while there is a slight improvement from $4$ to $8$ CSs as the global peak constraint prevents the algorithms from charging more EVs.}
 (iv) in integral revenue model, the proposed \iocs algorithm acts significantly better than \iolp. In particular, \iocs improves \iolp by \revv{$38\%, 36\%$, and $32\%$} for $m=2$, $m=4$ and $m=8$, respectively. \revv{In fractional revenue model, however, there is a very slight difference between \focs and \folp.}
 %We highlight that \iocs (\focs) is also better choice in terms of the algorithm complexity compared to \iolp (\folp).
(v) \ics approximates \textsc{iOpt} by \revv{$97\%$, $94\%$, and $95\%$} for $m=2$, $m=4$, and $m=8$, respectively. On average, \iocs is \revv{$90\%$} of \textsc{iOpt} and, \focs is \revv{$94\%$} of its optimal offline solution, \fcs.  
(vi) finally, the results depict that \ics achieves much better results in practice as compared to the theoretical approximation ratio that characterizes the performance in worst-case scenario.
